Graphical Tests for Power Comparison of Competing Designs

### 2012-03-25 intro.tex
intro to power

The strength of this evidence is determined by both the number of independent evaluations as well as the size of the lineup. 
Assume a lineup of size $m$ is evaluated by $n$ independent observers, out of which $x$ identify the data plot correctly, which we can use as an estimate for the probability to correctly identify the data plot. Correctly identifying the data plot also depends on the features visible in the other plots. 
We can think of a lineup as $m-1$ head-to-head comparisons of the dataplot and $m-1$ nullplots, which the dataplot all has to win, in order to be picked as the plot with the most distinct features.  This introduces the signal strength \cite{majumder:2011} of a particular lineup as:
\[
\text{signal strength} = 1- \left( \frac{x}{n}\right)^{1/(m-1)},
\]
where $m$ is the size of  the lineup, i.e. the number of individuals shown, and $x/n$ is the ratio of correctly identified data plots ($x$) in  $n$ independent evaluations. Signal strength is limited to a range of values between 0 and 1. Higher values indicate stronger signal.
 In the case of a parametric situation (i.e. the feature shown in the chart can be described in parametric form as a real-valued vector $\theta$ in the data), the signal strength is equivalent to the statistical power of the test. While we can compute 95\% confidence intervals for signal strength by using 95\% confidence intervals for $x/n$ given as $x/n \pm 1.96 \cdot \sqrt{x/n \cdot (1-x/n)}$ limited to the interval [0,1], the resulting interval is generally to large to be useful.

@@

>>>>

The strength of this evidence is determined by both the number of independent evaluations as well as the size of the lineup. 
Assume a lineup of size $m$ is evaluated by $n$ independent observers, out of which $x$ identify the data plot correctly, which we can use as an estimate for the probability to correctly identify the data plot. Correctly identifying the data plot also depends on the features visible in the other plots. 
We can think of a lineup as $m-1$ head-to-head comparisons of the dataplot and $m-1$ nullplots, which the dataplot all has to `win', in order to be picked as the plot with the most distinct features. Assuming that observers identify the plot with the stronger signal, the data plot is picked, if it has the strongest signal, leading us to a Beta distribution for the maximum.  This allows us to introduce the {\it signal strength} \cite{majumder:2011} of a lineup as:
\[
\text{signal strength} = 1- \left( \frac{x}{n}\right)^{1/(m-1)},
\]
where $m$ is the size of  the lineup, i.e. the number of individuals shown, and $x/n$ is the ratio of correctly identified data plots ($x$) in  $n$ independent evaluations. Signal strength is limited to a range of values between 0 and 1. Higher values indicate stronger signal.
 In the case of a parametric situation (i.e. the feature shown in the chart can be described in parametric form as a real-valued vector $\theta$ in the data), it can be shown \cite{majumder:2010} that the signal strength is approximately linear to the statistical power of the test. 
While we can compute 95\% confidence intervals for signal strength by using 95\% confidence intervals for $x/n$ given as $x/n \pm 1.96 \cdot \sqrt{x/n \cdot (1-x/n)}$ limited to the interval [0,1], the resulting interval is generally to large to be useful.

@@

The power of a statistical test is defined as the probability to reject the null hypothesis. Translated to the visual tests, power is the probability to identify the data plot from the lineup. While we can compute power only in the parametric setting of a simulation study, we can make use of the signal strength of a lineup.

Power consideration: what is statistical power, how can we assess it.
